---
title: "Practice SDM for Clubes de Ciencias!"
output: html_notebook
---

Here we go! Making our first SDM of clubes de ciencias, and doing it with computer programming in the R language!

```{r}
install.packages(c('raster', 'rgdal', 'dismo'))
```

### Trouble w/ getting `rJava`?
I was having some trouble getting `rJava` to load, so then I ran the following code:
source of this tip [here](https://zhiyzuo.github.io/installation-rJava/)
```{r}
dyn.load('/Library/Java/JavaVirtualMachines/jdk1.8.0_60.jdk/Contents/Home/jre/lib/server/libjvm.dylib')
install.packages("rJava")
library(rJava)
```


Now, import the set of occurrence data that we'll be using for our first SDM
```{r}
library(dismo)
file <- paste(system.file(package="dismo"), "/ex/bradypus.csv", sep="")
bradypus <- read.table(file,  header=TRUE,  sep=",")
# this is the file we will use:
```

Inspect the data
```{r}
bradypus <- read.table(file,  header=TRUE,  sep=",")
head(bradypus)
```

Since the data are all for the same species, we only need columns 2 and 3 for the longitude and latitue
```{r}
bradypus <- bradypus[,2:3]
head(bradypus)

```

