---
title: "Practice SDM for Clubes de Ciencias!"
output: html_notebook
---

Here we go! Making our first SDM of clubes de ciencias, and doing it with computer programming in the R language!

```{r}
# install.packages(c('raster', 'rgdal', 'dismo'))
```

### Trouble w/ getting `rJava`?
I was having some trouble getting `rJava` to load, so then I ran the following code:
source of this tip [here](https://zhiyzuo.github.io/installation-rJava/)
```{r}
dyn.load('/Library/Java/JavaVirtualMachines/jdk1.8.0_60.jdk/Contents/Home/jre/lib/server/libjvm.dylib')
# install.packages("rJava")
library(rJava)
```


Now, import the set of occurrence data that we'll be using for our first SDM
```{r}
library(dismo)
file <- paste(system.file(package="dismo"), "/ex/bradypus.csv", sep="")
bradypus <- read.table(file,  header=TRUE,  sep=",")
# this is the file we will use:
```

Inspect the data
```{r}
bradypus <- read.table(file,  header=TRUE,  sep=",")
head(bradypus)
```

Since the data are all for the same species, we only need columns 2 and 3 for the longitude and latitue
```{r}
bradypus <- bradypus[,2:3]
head(bradypus)
```

### The climate data
First, we've got to find all of the climate data
```{r}
path <- file.path(system.file(package="dismo"), 'ex')
files <- list.files(path, pattern='grd$', full.names=TRUE )
files
```

Then, we need to combine them all into a stack of different climate layers
```{r}
predictors <- stack(files)
predictors
```
Let's check out what they are called, and what they look like
```{r}
names(predictors)
plot(predictors)
```

### Mapping climate data and occurrence data
Now we can visually show, in the same window, one layer of climate data, plus a map of the world, plus the occurrence data for our species
```{r}
library(maptools)
data(wrld_simpl)
file <- paste(system.file(package="dismo"), "/ex/bradypus.csv", sep="")
bradypus <- read.table(file,  header=TRUE,  sep=',')
# we do not need the first column
bradypus  <- bradypus[,-1]
# first layer of the RasterStack
plot(predictors, 1)
# note the "add=TRUE" argument with plot
plot(wrld_simpl, add=TRUE)
# with the points function, "add" is implicit
points(bradypus, col='blue')
```

### Extracting data
Now, we extract climate information at every point where we have a GPS point.  This tells us about the type of climate that the species likes best.  Then after taking these steps we have a bunch of data points corresponding to the type of climate that the species likes best, and other data points corresponding to climate the invasive species doesn't like as much.
```{r}
set.seed(0)
group <- kfold(bradypus, 5)
pres_train <- bradypus[group != 1, ]
pres_test <- bradypus[group == 1, ]
```

#### background points
```{r}
ext <- extent(-90, -32, -33, 23) # Set your extent to the area we're sampling background from
pred_nf <- dropLayer(predictors, 'biome')
backg <- randomPoints(pred_nf, n=1000, ext=ext, extf = 1.25)
colnames(backg) = c('lon', 'lat')
group <- kfold(backg, 5)
backg_train <- backg[group != 1, ]
backg_test <- backg[group == 1, ]
```

### Maxent
Now, we make our predictive maps using the modeling software called `Maxent`
This part might take a little while, so be patient : )
```{r}
jar <- paste(system.file(package="dismo"), "/java/maxent.jar", sep='')
xm <- maxent(predictors, pres_train, factors='biome')
plot(xm)
```

Response plot
```{r}
response(xm)
```

Making our maps
```{r}
e <- evaluate(pres_test, backg_test, xm, predictors)
e
px <- predict(predictors, xm, ext=ext, progress='')
par(mfrow=c(1,2))
plot(px, main='Maxent, raw values')
plot(wrld_simpl, add=TRUE, border='dark grey')
tr <- threshold(e, 'spec_sens')
plot(px > tr, main='presence/absence')
plot(wrld_simpl, add=TRUE, border='dark grey')
points(pres_train, pch='+')

```

